---
title: HighFreq Package for High Frequency Time Series Data Management
author: Jerzy Pawlowski (algoquant)
abstract: Functions for chaining and joining time series, scrubbing bad data, managing time zones and alligning time indices, converting TAQ data to OHLC format, aggregating data to lower frequency, estimating volatility, skew, and higher moments.
tags: high frequency, time series, volatility
---

[![Build Status](https://travis-ci.org/algoquant/HighFreq.svg?branch=master)](https://travis-ci.org/algoquant/HighFreq)


### Overview

The motivation for the package *HighFreq* is to create a library of functions designed for managing trade and quote (*TAQ*) and *OHLC* data, and for efficiently estimating various statistics, like volatility, skew, Hurst exponent, and Sharpe ratio, from that data.  

The are several other packages which offer much of this functionality, like for example:

+ package [xts](https://cran.r-project.org/web/packages/xts/index.html)  

+ package [TTR](https://cran.r-project.org/web/packages/TTR/index.html)  

+ package [PerformanceAnalytics](https://cran.r-project.org/web/packages/PerformanceAnalytics/index.html)  

+ package [highfrequency](https://cran.r-project.org/web/packages/highfrequency/index.html)  

Unfortunately many of the functions in these packages are either too slow, or lack some critical functionality, or produce data in inconsistent formats (with *NA* values, etc.)  The package *HighFreq* aims to create a unified framework, with consistent data formats and naming conventions. 

The package *HighFreq* relies on *OHLC* price and volume data formatted as *xts* time series, because the *OHLC* data format provides an efficient way of compressing *TAQ* data, while preserving information about price levels, volatility (range), and trading volumes.  Most existing packages don't rely on *OHLC* data, so their statistical estimators are much less efficient than those in package *HighFreq*. 


### Running and Rolling Statistics Over Time Series Data

Definitions of running and rolling statistics (aggregations):

+ A statistic is some function of *OHLC* data, for example the difference between
*High* minus *Low* prices is a simple statistic.  Estimators of volatility, skew, and higher moments are also statistics.

+ The package *HighFreq* calculates two different types of statistics time series: *running* and *rolling* statistics.
The *running* statistics time series are calculated using individual bars of *OHLC* data, while *rolling* statistics time series are calculated using multiple bars of *OHLC* data. 

+ An example of a *running* statistic is a time series of squared differences of the *High* minus *Low* prices.  Each point in a *running* statistic depends only on a single bar of *OHLC* data (and possibly on the neighboring bars.) 

+ An example of a *rolling* statistic is a time series of average *Close* prices calculated over a rolling look-back window. 
Some *rolling* statistics can be calculated from *running* statistics by calculating weighted averages. 

+ The functions called *run_\** calculate *running* statistics based on each bar of *OHLC* data, and produce a single-column *xts* time series with the same number of rows as the *OHLC* time series.

+ The functions called *roll_\** calculate *rolling* statistics based on multiple bars of *OHLC* data taken from a rolling look-back window, and often produce a single-column *xts* time series with the same number of rows as the *OHLC* time series.

+ *Running* and *rolling* statistics also represent a form of data aggregation, and can include many standard technical indicators, like *VWAP*, *Bollinger Bands*, etc.

<br>


### Functions for data scrubbing, formatting, and aggregation

The package *HighFreq* contains several categories of functions designed for:

+ managing *TAQ* and *OHLC* time series, 

+ estimating running and rolling statistics over time series, 


The package *HighFreq* contains functions for:

+ chaining and joining time series, 

+ scrubbing bad data from time series, 

+ managing time zones and alligning time indices, 

+ converting *TAQ* data to *OHLC* format, 

+ aggregating data to lower frequency (periodicity), 

+ estimating running statistics from *OHLC* data, such as volatility, skew, and higher moments (functions called *run_\**), 

+ calculating rolling aggregations (VWAP, Hurst exponent, Sharpe ratio, etc.), 

+ calculating seasonality aggregations, 

+ creating random *TAQ* and *OHLC* time series, 


### Installation and loading

Install package *HighFreq* from github:  
```{r eval=FALSE}
install.packages("devtools")
devtools::install_github(repo="algoquant/HighFreq")
library(HighFreq)
```
<br>

Install package *HighFreq* from source on local drive:  
```{r eval=FALSE}
install.packages(pkgs="C:/Develop/R/HighFreq", repos=NULL, type="source")
# Install package from source on local drive using R CMD
R CMD INSTALL C:\Develop\R\HighFreq
library(HighFreq)
```
<br>

Build reference manual for package *HighFreq* from *.Rd* files:  
```{r eval=FALSE}
system("R CMD Rd2pdf C:/Develop/R/HighFreq")
R CMD Rd2pdf C:\Develop\R\HighFreq
```
<br>


### Data

Trade and Quote (*TAQ*) data contains intraday trades and quotes on exchange-traded stocks and futures.  *TAQ* data is spaced irregularly in time, with data recorded each time a new trade or quote arrives.  The rows of *TAQ* data contain the quoted and traded prices, and the corresponding quote size or trade volume.  

*TAQ* data can be aggregated into Open-High-Low-Close (*OHLC*) data.  *OHLC* data is evenly spaced in time, with each row containing the *Open*, *High*, *Low*, and *Close* prices, and the trade *Volume*, recorded over the past time interval (called a *bar* of data).  The *Open* and *Close* prices are the first and last trade prices recorded in the time bar.  The *High* and *Low* prices are the highest and lowest trade prices recorded in the time bar.  The *Volume* is the total trading volume recorded in the time bar.  

Aggregating *TAQ* data into *OHLC* data provides data compression, while preserving information about price levels, volatility (range), and trading volumes.  In addition, evenly spaced data allows analysis of multiple time series, since all the prices are given at the same moments of time.

The package *HighFreq* includes three *xts* time series called *SPY*, *TLT*, and *VXX*, containing intraday 1-minute *OHLC* data for the *SPY*, *TLT*, and *VXX* ETFs.  The package *HighFreq* also includes an *xts* time series called *SPY_TAQ* with a single day of *TAQ* data for the *SPY* ETF.  The data is set up for lazy loading, so it doesn't require calling `data(hf_data)` to load it before being able to call it.

The data source is the 
[Wharton Research Data Service](https://wrds-web.wharton.upenn.edu/wrds/)  

List all the data sets included in the *HighFreq* package:  
```{r eval=FALSE}
# list all datasets in package HighFreq
data(package="HighFreq")
```
<br>


### Examples

More examples can be found in the vignettes titled *managing_time_series* and *estimating_statistics*.  

Aggregate *TAQ* data into a 1-minute bar *OHLC* time series:  
```{r eval=FALSE}
# aggregate TAQ data to 1-min OHLC bar data, for a single symbol, and save to file
sym_bol <- "SPY"
save_scrub_agg(sym_bol, 
               data_dir="E:/mktdata/sec/", 
               output_dir="E:/output/data/")
```
<br>

Calculate daily trading volume:  
```{r eval=FALSE}
daily_volume <- apply.daily(x=Vo(SPY), FUN=sum)
colnames(daily_volume) <- "SPY.Volume")
chart_Series(x=daily_volume, name="daily trading volumes for SPY")
```
<br>

Calculate daily average open to close variance from minutely *OHLC* prices:  
```{r eval=FALSE}
# calculate daily average open to close variance
var_daily <- (6.5*60*60^2)*xts::apply.daily(x=SPY, FUN=agg_regate, 
                              mo_ment="run_variance", calc_method="rogers_satchell")
colnames(var_daily) <- "SPY.Var"
chart_Series(100*sqrt(var_daily["/2010"]), name="SPY daily standard deviation")
```
<br>

Calculate daily skew from minutely *OHLC* prices:  
```{r eval=FALSE}
skew_daily <- apply.daily(x=SPY, FUN=agg_regate, mo_ment="run_skew")
skew_daily <- skew_daily/(var_daily)^(1.5)
colnames(skew_daily) <- "SPY.Skew")
chart_Series(x=skew_daily, name="daily skew for SPY")
```
<br>

Calculate rolling prices:  
```{r eval=FALSE}
roll_prices <- rutils::roll_sum(Op(SPY), win_dow=10)/10
colnames(roll_prices) <- "SPY.Rets"
# plot candle chart
chart_Series(SPY["2013-11-12", ], name="SPY Prices")
add_TA(roll_prices["2013-11-12"], on=1, col="red", lwd=2)
```
<br>

Calculate rolling volume-weighted variance:  
```{r eval=FALSE}
var_rolling <- roll_moment(oh_lc=SPY["2012"], mo_ment="run_variance", win_dow = 10)
# plot without overnight jump
chart_Series(var_rolling["2012-11-12", ][-(1:11)], name="SPY rolling volume-weighted variance")
```
<br>

Calculate daily seasonality of variance:  
```{r eval=FALSE}
var_seasonal <- season_ality((24*60*60^2)*run_variance(oh_lc=SPY))
colnames(var_seasonal) <- "SPY.var_seasonal"
chart_Series(x=var_seasonal, name="SPY variance daily seasonality")
```
