---
title: Estimating Running and Rolling Statistics Over Time Series
author: Jerzy Pawlowski (algoquant)
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"HighFreq Package"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Below are examples of applying functions from the package *HighFreq* for estimating running and rolling statistics over time series


Calculate rolling average closing prices using package *rutils*:
```{r eval=FALSE}
# load HighFreq to load SPY data
library(HighFreq)
# rolling average prices
win_dow <- 10
prices_rolling <- rutils::roll_sum(Cl(SPY), win_dow=win_dow)/win_dow
colnames(prices_rolling) <- "SPY.Prices"
chart_Series(SPY["2013-11-12"], name="SPY Prices")
add_TA(prices_rolling["2013-11-12"], on=1, col="red", lwd=2)
legend("top", legend=c("SPY prices", "average prices"), 
       bg="white", lty=c(1, 1), lwd=c(2, 2), 
       col=c("black", "red"), bty="n")
```
<br>


Calculate rolling volume-weighted average closing prices using package *HighFreq*:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# rolling volume-weighted average prices
win_dow <- 10
prices_rolling <- roll_vwap(oh_lc=SPY["2013-11-12"], win_dow=win_dow)
colnames(prices_rolling) <- "SPY.Prices"
chart_Series(SPY["2013-11-12"], name="SPY VWAP Prices")
add_TA(prices_rolling["2013-11-12"], on=1, col="red", lwd=2)
legend("top", legend=c("SPY prices", "VWAP prices"), 
       bg="white", lty=c(1, 1), lwd=c(2, 2), 
       col=c("black", "red"), bty="n")
```
<br>


Calculate variance from *OHLC* prices using different range estimation methods:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# calculate variance of SPY using method yang_zhang
# scale from minutely to daily frequency and also apply factor to compensate for secondly units
vari_ance <- (6.5*60*60^2)*HighFreq::calc_variance(SPY, calc_method="yang_zhang")
# calculate variance of SPY without accounting for overnight jumps
vari_ance <- (6.5*60*60^2)*HighFreq::calc_variance(SPY, calc_method="rogers_satchell")
```
<br>


Calculate rolling variance and skew from minutely *OHLC* prices:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# calculate running variance using method rogers_satchell
# scale from minutely to daily frequency and also apply factor to compensate for secondly units
var_running <- (6.5*60*60^2)*run_variance(oh_lc=SPY, 
                                          calc_method="rogers_satchell")
# calculate rolling volume-weighted average daily variance
win_dow <- 21
var_rolling <- roll_vwap(oh_lc=SPY, x_ts=var_running, win_dow=win_dow)
colnames(var_rolling) <- "SPY.Var"

# calculate rolling daily variance using roll_variance()
var_rolling <- (6.5*60*60^2)*roll_variance(oh_lc=SPY, 
                                          calc_method="rogers_satchell", 
                                          win_dow=win_dow)

# calculate rolling volume-weighted average skew indicator
skew_running <- run_skew(oh_lc=SPY)
skew_rolling <- roll_vwap(oh_lc=SPY, x_ts=skew_running, win_dow=win_dow)
skew_rolling <- ifelse(var_rolling>0, skew_rolling/(var_rolling)^(1.5), 0)
colnames(skew_rolling) <- "SPY.Skew"
chart_Series(skew_rolling["2013-11-12"], name="SPY Rolling Skew")
```
<br>


Calculate rolling variance and skew using function *roll_moment()* (produces same result as above):
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# calculate rolling volume-weighted average variance and skew
win_dow <- 21
var_rolling <- roll_moment(oh_lc=SPY, win_dow=win_dow)
skew_rolling <- roll_moment(oh_lc=SPY, mo_ment="run_skew", win_dow=win_dow)
skew_rolling <- ifelse(var_rolling>0, skew_rolling/(var_rolling)^(1.5), 0)
chart_Series(skew_rolling["2013-11-12"], name="SPY Rolling Skew")
```
<br>


Calculate daily average open to close variance and skew from minutely *OHLC* prices:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# calculate daily average open to close variance
var_daily <- (6.5*60*60^2)*xts::apply.daily(x=SPY, FUN=agg_regate, 
                              mo_ment="run_variance", calc_method="rogers_satchell")
colnames(var_daily) <- "SPY.Var"
chart_Series(100*sqrt(var_daily["/2010"]), name="SPY daily standard deviation")

# calculate daily average skew
skew_daily <- xts::apply.daily(x=SPY, FUN=agg_regate, mo_ment="run_skew")
skew_daily <- skew_daily/(var_daily)^(1.5)
colnames(skew_daily) <- "SPY.Skew"
inter_val <- "2013-10/2013-12"
chart_Series(skew_daily[inter_val], name="SPY Skew")
```
<br>


Calculate intraday seasonality of trading volumes and variance:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
volume_seasonal <- season_ality(Vo(SPY))
colnames(volume_seasonal) <- "SPY.volume_seasonal"
chart_Series(volume_seasonal, name="SPY intraday seasonality of volume")
var_seasonal <- season_ality((6.5*60*60^2)*run_variance(oh_lc=SPY))
colnames(var_seasonal) <- "SPY.var_seasonal"
chart_Series(var_seasonal, name="SPY intraday seasonality of variance")
```
<br>


Apply an aggregation function over a rolling lookback window and the end points of an *OHLC* time series:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# extract a single day of SPY data
x_ts <- SPY["2012-02-13"]
win_dow <- 11
# calculate the rolling sums of the columns of x_ts
agg_regations <- roll_apply(x_ts, agg_fun=sum, win_dow=win_dow, by_columns=TRUE)
# define a vector-valued aggregation function
agg_function <- function(x_ts)  c(max(x_ts[, 2]), min(x_ts[, 3]))
# apply the aggregation function over a rolling window
agg_regations <- roll_apply(x_ts, agg_fun=agg_function, win_dow=win_dow)
# define end points at 11-minute intervals (SPY is minutely bars)
end_points <- rutils::end_points(x_ts, inter_val=win_dow)
# calculate the rolling sums of the columns of x_ts over end_points
agg_regations <- roll_apply(x_ts, agg_fun=sum, win_dow=2, end_points=end_points, by_columns=TRUE)
# apply the vector-valued aggregation function over the end_points of x_ts
agg_regations <- roll_apply(x_ts, agg_fun=agg_function, win_dow=2, end_points=end_points)
```
<br>


