---
title: Managing *TAQ* and *OHLC* Time Series
author: Jerzy Pawlowski (algoquant)
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"HighFreq Package"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Below are examples of applying functions from the package *HighFreq* for managing *TAQ* and *OHLC* time series


Scrub a single day of *TAQ* data contained in *SPY_TAQ* "by hand", without aggregating it to *OHLC*:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# create local copy of SPY TAQ data
ta_q <- SPY_TAQ
# convert timezone of index to New_York
index(ta_q) <- lubridate::with_tz(time=index(ta_q), tzone=tzone)
# subset data to NYSE trading hours
ta_q <- ta_q['T09:30:00/T16:00:00', ]
# remove duplicate time stamps using duplicated()
ta_q <- ta_q[!duplicated(index(ta_q)), ]
# scrub quotes with suspect bid-offer spreads
bid_offer <- ta_q[, "Ask.Price"] - ta_q[, "Bid.Price"]
sus_pect <- extreme_values(bid_offer, win_dow=win_dow, vol_mult=vol_mult)
# remove suspect values
ta_q <- ta_q[!sus_pect]
# calculate mid prices
mid_prices <- 0.5 * (ta_q[, "Bid.Price"] + ta_q[, "Ask.Price"])
# replace whole rows containing suspect price jumps with NA, and perform locf()
ta_q[price_jumps(mid_prices, win_dow=31, vol_mult=1.0), ] <- NA
ta_q <- na.locf(ta_q)
```
<br>


Scrub and aggregate a single day of *TAQ* data to *OHLC* format using function scrub_agg():
```{r eval=FALSE}
# scrub and aggregate a single day of SPY TAQ data to OHLC
oh_lc <- HighFreq::scrub_agg(ta_q=SPY_TAQ)
chart_Series(oh_lc, name=sym_bol)
```
<br>


*TAQ* data is organized in individual daily files. The function *save_scrub_agg()* reads all the *TAQ* files for a single symbol, and converts them into a single *OHLC* file. 
For a single symbol, aggregate all the daily *TAQ* data to 15-min *OHLC* bar data, and save the data to a single file:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# set data directories
data_dir <- "C:/Develop/data/hfreq/src/"
output_dir <- "C:/Develop/data/hfreq/scrub/"
sym_bol <- "SPY"
# aggregate SPY TAQ data to 15-min OHLC bar data, and save the data to a file
save_scrub_agg(sym_bol=sym_bol, data_dir=data_dir, output_dir=output_dir, period="15 min")
```
<br>


For a list of symbols, aggregate all the daily *TAQ* data to 15-min *OHLC* bar data (one symbol at a time), and save the data to a individual files:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
sym_bols <- c("SPY", "TLT", "VXX")
# aggregate TAQ data to 15-min OHLC bar data, and save the data to a file
sapply(sym_bols, save_scrub_agg, 
       data_dir=data_dir, output_dir=output_dir, period="15 min")
```
<br>


For a single symbol, load a single day of *TAQ* data, scrub it using function scrub_taq(), and save the scrubbed TAQ data to a single *.RData* file:
```{r eval=FALSE}
# load HighFreq
library(HighFreq)
# set data directories
data_dir <- "C:/Develop/data/hfreq/src/"
output_dir <- "C:/Develop/data/hfreq/scrub/"
sym_bol <- "SPY"
# load a single day of TAQ data
file_name <- paste0(sym_bol, "/2014.05.02.", sym_bol, ".RData")
sym_bol <- load(file.path(data_dir, paste0(sym_bol, "/2014.05.02.", sym_bol, ".RData")))
# scrub a single day of TAQ data, without aggregating it to OHLC
ta_q <- scrub_taq(ta_q=get(sym_bol))
save(list=sym_bol, file=file.path(output_dir, file_name))
```
<br>
